Modeling and leveraging intuitive theories to improve vaccine attitudes

### 2022-07-08 paper/paper-papaja.Rmd
paper revisions

note: |
  This manuscript has not yet been peer-reviewed (2/15/2022).

authornote: |
  All data, and analysis code for this project are available at https://osf.io/sdfpj/. Preregistration information for Studies 3a and 3b is available at https://osf.io/csnez/registrations. Earlier modeling and analyses of data from Study 1 were presented at the 40th Annual Conference of the Cognitive Science Society. A six-page paper also appeared in the proceedings [@powell.etal2018].

@@

The difficulty of convincing people to vaccinate themselves and their children illustrates a general pattern of findings showing that people often do not respond to evidence in ways that educators might wish [e.g. @lewandowsky.etal2012]. Part of the challenge in correcting misconceptions is that they are situated within larger, internally coherent belief systems---intuitive theories---that guide how people interpret and respond to evidence [e.g., @gershman2019; @lewandowsky.etal2012; @jern.etal2014; @cook.lewandowsky2016].

@@

We argue that correcting misconceptions requires presenting evidence in a way that is sensitive to and persuasive within the broader conceptual context in which those misconceptions reside [see @gershman2019; @weisman.markman2017; @cook.lewandowsky2016]. Understanding the revision of any one specific belief (e.g., a misconception like *vaccines are dangerous*) demands a holistic understanding of the wider set of interrelated beliefs that make up people's intuitive theories.

@@

Quine argued that all of one’s beliefs are intertwined in a “web of belief” [@quine1951; @quine.ullian1978]---an apt metaphor for intuitive theories. Intuitive theories consist of relations between beliefs that specify how states of affairs causally, logically, or otherwise interconnect with one another. Given this definition, we argue that any set of beliefs related by an intuitive theory can be expected to exhibit several qualitative features:

1) Where beliefs about two states of affairs, A and B, are related through a shared intuitive theory (e.g., A causes B), those beliefs will be systematically correlated across individuals (e.g. people who believe A will also tend to believe B);

2) When two beliefs are related by an intuitive theory, evidence affecting one of those beliefs will also affect the other in accordance with their relationship in the intuitive theory (e.g. evidence for B will increase credence in A, and vice versa).

3) Thus, across individuals, the average change in one belief following evidence affecting another belief will be proportional to the correlation between those beliefs across individuals (e.g. evidence for B will increase credence in A in proportion to the correlation between beliefs A and B).

Here, we sought to formalize this understanding of intuitive theories by representing an intuitive theory as a Bayesian network. Originally designed as tools for artificial intelligence [@pearl1988], Bayesian networks are a class of probabilistic graphical models that have been deployed widely across cognitive science [e.g. @gopnik.etal2004; @jern.etal2014; @griffiths.tenenbaum2005]. The heart of a Bayesian network is a graphical structure ---not unlike a "web of belief"---that represents variables as nodes connected by edges, where those edges represent probabilistic or statistical relations.

For example, the graph $\text{FIRE} \rightarrow \text{SMOKE}$ represents a Bayesian network capturing the understanding  that fire causes smoke. The example of cause and effect is especially intuitive, but edges in a Bayesian network need not be causal; they are better thought of as representing “influence relations,” which may also include relationships like logical implication and set membership [@williamson2001]. This influence relationship is represented as a probabilistic function, where the probability of smoke depends on the probability of fire.

This model captures each of the qualitative features laid out above: First, it captures people’s  reasoning  about these states of affairs given observations: observing smoke would lead someone to infer that there is likely to be a fire, and observing fire would lead someone to predict smoke will soon follow (Feature 1). This model can also capture how these beliefs might covary across individuals: if two people hold different degrees of belief in FIRE, this model specifies the degree of belief each should have in SMOKE (Feature 2). Because the conditional probability of SMOKE given FIRE determines both the change in beliefs given observations and the correlation of beliefs across individuals, changes in beliefs following evidence will be proportional to their correlations across individuals (Feature 3). Formulating these features in a formal model, rather than a verbal theory, allows us to specify not just the presence but the quantitative strength of these relations, and to simulate the cumulative influences of many factors in more complex networks (for example, incorporating beliefs about the frequency of wildfires, recent weather conditions, etc.).

@@

Fortunately, the formalism of Bayesian networks provides tools for navigating these challenges. Roughly, when the relationships among beliefs are unknown, we can use the qualitative features of intuitive theories to *infer* the structure of an intuitive theory from observed correlations among beliefs. Structure learning algorithms [see @scutari.denis2021] provide a means for searching through the (super-exponentially large) possible space of Bayesian networks to find the structure most likely to have generated some observed data. These approaches allow us to scale the Bayesian network modeling framework beyond the smaller models characteristic of early computational work on intuitive theories [e.g., @cook.lewandowsky2016; @griffiths.tenenbaum2005; @jern.etal2012].

@@

Across all our studies, our primary outcome of interest was participants’ intentions to vaccinate their children against diseases like measles, mumps, and rubella (which we refer to henceforth as *vaccination intentions*). Drawing on a variety of sources, including research on anti-vaccine skepticism [e.g.,  @dube.etal2015; @larson.etal2015; @salmon.etal2015; @williams2014], anti-vaccine websites, and our own qualitative work with vaccine skeptics (see SI), we generated a list of 13 additional beliefs that, though surely not exhaustive, might jointly offer a reasonable means of understanding and predicting vaccine intentions.

@@

Each scale measured agreement with 4-6 statements, including at least one reverse-coded item, and was highly reliable (observed reliability, Study 1: all Cronbach’s $\alpha \geq .73$). We developed 13 of these scales, and included one preexisting scale (the “holistic balance” subscale from [@mcfadden.etal2010]. See SI for all items included in all scales.

In Study 1, we administered these 14 belief scales to participants and in an online survey with `r results$s1$demo$n_include` participants. We then used these intercorrelations to build a cognitive model of the relations among these beliefs via Bayesian network structure learning techniques.

@@

A total of `r results$s1$demo$n_total` people recruited from Amazon Mechanical Turk (mTurk) participated in this study conducted in August of 2017. Participants were paid \$1.60 for about 8 minutes of their time. `r results$s1$demo$n_exclude` participants (6%) failed at least one of two attention check questions and were excluded from further analysis, leaving a final sample of n = `r results$s1$demo$n_include`. See SI for demographic information about this sample.

@@

The goal of this study was to use these data to conduct a bottom-up search for a Bayesian network that would represent the intuitive theory surrounding vaccination decisions as a network of related beliefs. There were strong intercorrelations among these 14 beliefs confirming that these beliefs are, in fact, closely related to each other and to intentions to vaccinate children (see SI). Based on the observed associations among participants' beliefs about each of theses issues, we sought to infer how these states of affairs (e.g., that vaccines are dangerous, that diseases are severe) are connected in people's intuitive theories---how people understand these states of affairs to cause, imply, or otherwise influence one another.

@@

A Bayesian network is a directed acyclic graph that represents variables as nodes connected by directed edges. Each set of “child” and “parent” nodes in the larger network is a “family”. The network’s directed edges encode conditional dependencies among variables, and many parameterizations for the individual families are possible. In our model, the probability of some “child” belief node $y$ can be expressed as a noisy-logical function of the influence of one or more “parent” belief nodes $x$:

$$P(y=1) = \Big(1-(1-w_0)\prod_i(1-w_i)^{x_i}\Big)\prod_j(1-w_j)^{x_j}$$

@@

A cognitive model composed of these types of relations is capable of capturing a causal system of non-interactive generative and preventive causes. But it bears noting that this simplification comes with limitations. For instance, returning to the example network in Figure 1d, this parameterization would not permit capturing the relationships between dryness, lightning, and fire. Dryness is an enabling condition that both increases the probability of fire but also increases the probability that lightning would cause a fire. So the influences of these variables are not independent and would need a more complex parameterization to be properly represented.

Finally, having transformed participants’ scores on each belief scale to fall between 0 and 1, we estimated the parameters of this model as a Beta regression using maximum likelihood methods (see SI for details).

@@

We aimed to represent not only the relationships among variables in the data but the *cognitive model* that gave rise to these observed relationships, i.e., the relationships among beliefs in the mind. We actually estimated two models: First, we estimated a data model assuming that participants possess a shared mental model from which they generate their responses to the belief scales. In this model, the $x$ in the equation above corresponded to participants’ continuous responses to the belief scales (e.g. a participant's score for the *vaccine danger* scale). Then, from the parameters estimated in the data model we created a cognitive model of participants’ reasoning, where the $x$ are binary, indicating whether that state of affairs obtains (e.g., whether “vaccines are dangerous”). We used this cognitive model to simulate how beliefs could change following exposure to evidence.

@@

To conduct an efficient search through the super-exponential space of graphs, we tested several learning algorithms implemented in the “bnlearn” R package [@scutari2010] and evaluated their performance through cross validation in our training split (see SI). Following this analysis, we chose to fit our model with the score-based “hill-climbing” algorithm, scoring models via the Bayesian Information Criteria, an asymptotic estimate of the likelihood of the data under a given graph structure. This produced the network shown in Figure 1. See SI for more detailed explanations of these modeling choices.

#### Structure learning

To avoid overfitting, we used an 80% split of the data (*n* = 904) for training, and reserved the remaining 20% for evaluating model performance. We compared performance of different structure learning algorithms using k-fold cross validation within the training split. Based on these results, we chose to apply a hill-climbing learning algorithm using our generative model constraints (see SI for details). Applying these techniques resulted in an estimate of the network shown in Figure 1.

```{r fig1-maingraph, fig.align='center', out.width="90%", fig.cap="Bayesian network cognitive model resulting from application of hill-climbing algorithm with our generative-model constraints applied to data in Study 1 training split (n = 904). The network has 29 edges connecting these 14 variables. Edge polarity and weights are indicated by color, with red hues indicating negative or preventive influences, and blue hues indicating positive or generating influences. The direction of the edge was not determinable from the data and modeling constraints for two edges: $\\text{Naturalism} \\rightarrow \\text{Holistic Balance}$ and $\\text{Medical Skepticism} \\rightarrow \\text{Parental Expertise}$; see SI."}

@@

We evaluated the resulting cognitive model’s performance by testing its accuracy in predicting responses in the 20% testing split (*n* = 226). We generated predictions for held-out participants’ responses for each variable by conditioning the network on the remaining 13 observed variables. Figure 2 compares the model’s predictions with participants’ actual responses. Collapsing across all variables, the average correlation between predicted and observed responses was *r* = `r round(avg_corr, 3)`, accounting for `r round(100*avg_corr^2, 0)`% of the variance in observed responses.

Our goal in developing this cognitive model was to create a psychologically plausible and intelligible model of the intuitive theory surrounding vaccination decisions. However, it is also reasonable to ask whether other less-constrained models might be more predictive. To address this, we compared the Bayesian network model’s predictions to a suite of 14 naive “maximal” regression models, each using values on all 13 other belief scales as predictors for each belief. Our simpler model provided equally strong predictions (see SI). And, as will be evidenced in later studies, our model can also be used to generate accurate predictions about how beliefs will change following evidence.

```{r fig2-validation, fig.align='center', out.width="100%", fig.cap="Observed responses and cognitive model predictions for each belief among participants in the Study 1 testing split. Correlations between observed and predicted values ranged from .380 to .898 across the different belief scales. In general, the model showed greater predictive accuracy for more central beliefs (e.g., \\emph{vaccine danger}) than for more peripheral beliefs (e.g., \\emph{parental protectiveness}; see SI for details)."}
knitr::include_graphics("../local/valid_plt.png")
```

What does this cognitive model reveal about the intuitive theory underlying vaccination decisions?

@@

It is worth pointing out that not all of the relationships represented in the model should be interpreted as _causal_. Due to Bayesian networks’ tight connections with structural causal networks, it can be tempting to give the edges this interpretation, but in some cases these relationships are better thought of as probabilistic dependencies or “influence” relations (Williamson, 2001). For instance, intuitively we might imagine that vaccines having toxins (_toxic additives in vaccines_) could cause them to be dangerous (_vaccine danger_). These nodes are connected in the model, but the edges between them run in the opposite direction, and therefore do not support a clear causal interpretation. One interpretation is that this edge instead reflects a set membership relation; that toxins are one type of danger vaccines might present.

In other cases, our data and explicit constraints do not perfectly identify the directions of the edges. This is the case for two edges (Naturalism $\rightarrow$  Holistic Balance and Medical Skepticism $\rightarrow$ Parental Expertise; see SI for details). In both cases, we do not feel it is intuitive which direction the edge should flow in, nor even that it should be directed at all. Such a counterintuitive result might be taken as indicating a limitation of Bayesian network models, which are composed solely of directed, acyclic relationships. Prior work has successfully modeled coherent belief revision by leveraging models with undirected or bidirected relations [e.g. @spellman.etal1993]. Alternatively, we might imagine there is an additional unmeasured variable that connects these beliefs that could more intelligibly account for their association within the context of a directed network. For instance, a more abstract epistemic belief might influence both medical skepticism and parental expertise beliefs. To speculate, such an abstract belief might be in one's own expertise or in how we come to have knowledge (e.g. through direct experience versus science).

@@

Very few educational interventions have been shown to improve attitudes toward childhood vaccination. One of the few successful educational interventions in this domain was tested by Horne and colleagues [-@horne.etal2015], who found that emphasizing the dangers of preventable diseases like mumps, measles, and rubella improved participants’ attitudes toward vaccines. Their intervention included descriptions of the symptoms of mumps, measles, and rubella; vivid photos of children suffering from these diseases; and a brief story from the perspective of a parent whose child was hospitalized with measles (see SI). This “disease risk” intervention improved attitudes toward vaccines compared to an irrelevant control essay and compared to an intervention aimed at debunking concerns about vaccines causing autism. In Study 2, we aimed to replicate this success and leverage the cognitive model developed in Study 1 to gain insight into how and why such an intervention successfully induces  belief revision in this domain.

In Study 2, we used the belief network established in Study 1 to predict how Horne and colleagues’ [-@horne.etal2015] intervention would affect the beliefs in this network, including people’s intentions to vaccinate their children. Participants completed the 14 belief scales from Study 1, then returned two days later to complete either the disease risk intervention or a no-intervention control, followed by the same 14 belief scales.

@@

Participants (*n* = `r results$s2$demo$n_complete` passed attention checks in total) in the intervention conditions were presented with a version of Horne and colleagues'  “disease risk” intervention, modified to present subparts in a fixed (rather than randomized) order, with slight changes to wording to improve clarity, and with a final statement summarizing the take-home message of the intervention; see SI. They then completed the same 14 belief scales and demographic questions that they had completed in Phase 1.

@@

Figure 3 shows the change in participants’ attitudes across all 14 belief scales, from pretest to posttest. As expected, the largest change was observed for *disease severity* beliefs---the belief directly targeted by the intervention. Most importantly, the disease risk intervention significantly increased participant’s *vaccination intentions*, conceptually replicating earlier findings regarding this intervention [@horne.etal2015].

@@

To model the effects of the intervention presented in Study 2 on the intuitive theory surfaced in Study 1, we augmented the belief network representing the intuitive theory with an additional node representing the information about measles, mumps, and rubella presented in the intervention. This evidence node was added as a child of the *disease severity* node, allowing the intervention to act as virtual evidence for *disease severity* [following @pearl1988].

@@

$$logit(posttest) = logit(pretest) + \ln(ER_0) + x\ln(ER_{1})$$

@@

$$ER_i = \ln\frac{(P(d_i|h=1))}{(P(d_i|h=0))}$$

>>>>

# note: |

authornote: |
  Supplementary Materials, including all data and analysis code for this project are available at https://osf.io/sdfpj/. Preregistration information for Studies 3a and 3b is available at https://osf.io/csnez/registrations. Earlier modeling and analyses of data from Study 1 were presented at the 40th Annual Conference of the Cognitive Science Society. A six-page paper also appeared in the proceedings [@powell.etal2018].

@@

The difficulty of convincing people to vaccinate themselves and their children illustrates a general pattern of findings showing that people often do not respond to evidence in ways that educators might wish [e.g. @lewandowsky.etal2012]. Part of the challenge in correcting misconceptions is that they are situated within larger, internally coherent belief systems---intuitive theories---that guide how people interpret and respond to evidence [e.g., @cook.lewandowsky2016; @gershman2019; @lewandowsky.etal2012; @jern.etal2014]. In particular, prior work suggests the "attitude roots" of vaccine skepticism may spread quite wide, with vaccine misconceptions correlated with quite distant beliefs about the world [@hornsey.etal2018].

@@

We argue that correcting misconceptions requires presenting evidence in a way that is sensitive to and persuasive within the broader conceptual context in which those misconceptions reside [see @cook.lewandowsky2016; @hornsey.etal2018; @gershman2019; @weisman.markman2017]. Understanding the revision of any one specific belief (e.g., a misconception like *vaccines are dangerous*) demands a holistic understanding of the wider set of interrelated beliefs that make up people's intuitive theories.

@@

Quine argued that all of one’s beliefs are intertwined in a “web of belief” [@quine1951; @quine.ullian1978]---an apt metaphor for intuitive theories. Intuitive theories consist of relations between beliefs that specify how states of affairs causally, logically, or otherwise interconnect with one another. Specifically, we consider intuitive theories to be generative mental models that entail probabilistic dependencies among possible states of affairs.
Given this definition, we argue that any pair of beliefs related by a intuitive theory can be expected to exhibit several qualitative features:

1. When two beliefs are related by an intuitive theory, evidence directly affecting one of those beliefs will also affect the other in accordance with their relationship in the intuitive theory (e.g. evidence for B will increase credence in A, and vice versa). This is a clear and direct consequence of the laws of probability and Bayes' rule. 
2. Where beliefs about two states of affairs, A and B, are related through an intuitive theory (e.g., A causes B), those beliefs will be systematically correlated across individuals in proportion to the average conditional probabilities (e.g. people who believe A will also tend to believe B); This can be seen through an application of the law of total probability, which implies a linear relationship between the two beliefs across individuals. 
3. Thus, we can often reasonably expect there will be some systematic relationship between the average change in each belief following observation of some evidence and the correlation among those beliefs across individuals. We provide a more formal explication of these last two points in Supplementary Materials.

Here, we sought to formalize this understanding of intuitive theories by representing an intuitive theory as a Bayesian network. Originally designed as tools for artificial intelligence [@pearl1988], Bayesian networks are a class of probabilistic graphical models that have been deployed widely across cognitive science [e.g. @gopnik.etal2004; @jern.etal2014; @griffiths.tenenbaum2005]. The heart of a Bayesian network is a graphical structure---not unlike a "web of belief"---that represents variables as nodes connected by edges, where those edges represent probabilistic or statistical relations.

### Bayesian networks

Figure 1 presents several examples of Bayesian networks. Each of the networks is a _directed acyclic graph_ (DAG), a type of graphical model. Graphical models represent variables as nodes connected by edges. In a DAG, these edges are directed, flowing from parent to child, and are constrained so that they cannot form cycles or loops. In a Bayesian network, these directed edges encode conditional dependencies among variables, so that the conditional probability distribution for a given variable is defined only in terms of its direct parents.

(ref:bn-explain-caption) Several examples illustrating different features of Bayesian Networks. \newline A) A graphical representation of influence relations: Fire causes smoke. This influence is represented by the conditional dependencies captured in a conditional probability table (CPT). \newline B) A “parent” node can be linked to multiple “children” and a “child” node can be linked to multiple “parents”: A child and its parents together compose a “family”. This graph has three families: {Accidents, lightning, drought, fire}, {fire, smoke}, and {fire, ash}. \newline C) Not all influence relations correspond to causation: we could say a storm _consists_ of rain and lightning (marked as dashed arrows), which in turn causes thunder. Causes can be generative or preventive, e.g. rain prevents drought and fire (red arrows). Some nodes may exert countervailing influences on others: A storm consists of rain, which can prevent fires, but a storm also consists of lightning, which can cause fires. In addition, parents may influence their children independently or interactively. Lightning and accidents could be seen as independent causes of fire—an accident may cause a fire whether or not there is lightning, and vice versa. However, drought conditions interact with both of these other parent nodes, making them more likely to produce a fire. In the present work, we will consider only the role of independent influence relations.

```{r, out.width='100%', fig.align='center', fig.cap = "(ref:bn-explain-caption)"}
knitr::include_graphics("fig-bn-examples.png")
```

In Figure 1a, the edge from $fire \rightarrow smoke$ indicates that the probability of smoke is increased by fire. So, conditional on fire, smoke becomes more likely. The Bayesian network captures how conditioning on any one node affects the probability of all the other nodes in the network. The directed edges in a Bayesian network capture _asymmetric_ relationships: different edge directions imply different conditional dependencies and independencies. A criterion called “d-separation” provides a formal means for deciding which nodes in a network conditionally depend on one another [@pearl1988]. To illustrate, consider some relationships depicted in Figure 1c. Intuitively, it should be clear that observing (conditioning on) lightning will affect the probability of fire and all of its descendents. Indeed, these nodes are “d-connected” in the network. But, observing (conditioning on) lightning should not affect the probability of accidents (these nodes are d-separated in the network).

It can be tempting to interpret the edges in a Bayesian network as representing causal relationships. Strictly speaking however, the Bayesian network is simply an encoding of the probabilistic relationships among variables. These dependencies can be interpreted most generally as representing “influence relations” [@williamson2001]. As modelers, we might sometimes give these influence relationships more specific interpretations: causal relationships and logical implications [@williamson2001] are two prominent sorts of influence relations. In other cases, influence relations might be interpreted as representing “constituency” (e.g. see Figure 1c) or an entity's having a particular property (e.g. an edge from "being an apple" to "being red" might represent that redness is a common property of apples). However, in all these cases these interpretations are external to the Bayesian network model itself, where the edges signify only probabilistic relationships.

Bayesian networks like those depicted in Figure 1 can be made to capture each of the qualitative features of intuitive theories laid out above. First, a Bayesian network captures people’s reasoning about these states of affairs given observations: observing smoke would lead someone to infer that there is likely to be a fire, and observing fire would lead someone to predict smoke will soon follow (Feature 1). Using the conditional probability values in Figure 1a, observing smoke would shift someone with a prior probability of fire $p(fire) = .2$ to a posterior probability of fire $p(fire|smoke) = .83$. And for this same person, observing fire would shift their credence in smoke from $p(smoke) = .238$ to $p(smoke|fire) = .99$.

This model can also capture how these beliefs might covary across individuals: if two people hold different degrees of belief in _fire_, but share the same understanding of the relationship between _fire_ and _smoke_, this model specifies the degree of belief each should have in _smoke_ (Feature 2). If Alice holds that $p(fire) = .4$ and Bob holds that $p(fire) = .1$, then Alice should belief $p(smoke) = (.99)(.4) + (.05)(.6) = .426$ and Bob should believe $p(smoke) = (.99)(.1) + (.05)(.9) = .144$.

We can leverage both of these facts to derive a model of people’s intuitive theory relating these affairs based on the correlations among these beliefs. Because the conditional probabilities of _smoke_ given _fire_ and $\neg fire$ crucially determine both the change in beliefs given observations and the correlation of beliefs across individuals, average changes in beliefs following evidence will generally be related to the initial correlation of those beliefs across individuals (Feature 3; see Supplementary Materials for further explanation).

Formulating these features in a formal model, rather than a verbal theory, allows us to specify not just the presence but the quantitative strength of these relations, and to simulate the cumulative influences of many factors in more complex networks (for example, incorporating beliefs about the frequency of wildfires, recent weather conditions, etc.).

Other network-based approaches have also been applied to model systematic relationships among beliefs and attitudes, such as neural network models [e.g. see @holyoak.powell2016; @holyoak.simon1999; @simon.etal2004; @thagard1989; @spellman.etal1993]. In these models, beliefs or attitudes are represented by nodes (neurons) connected by undirected or bi-directed edges—in contrast to the directed edges used in Bayesian Networks. One application of neural-network models has been the modeling of “coherence-shifts”: patterns of belief revision wherein the beliefs or attitudes that should putatively serve as inputs to an evaluation are themselves modified during the process of evaluation. However, the directed edges of Bayesian networks are similarly capable of capturing these coherent reasoning dynamics: When conditioning on a node within a network, information flows in both directions along the edges. To illustrate, consider how observing smoke will lead a reasoner to infer that there is a fire, just as readily as observing fire will lead them to predict there will be smoke.[^1]

[^1]: With this in mind, Bayesian Networks also offer potential models of coherence-based reasoning effects based on rational belief updating mechanisms [@jern.etal2014].

@@

Fortunately, the formalism of Bayesian networks provides tools for navigating these challenges. Roughly, when the relationships among beliefs are unknown, we can use the qualitative features of intuitive theories to *infer* the structure of an intuitive theory from observed correlations among beliefs. Structure learning algorithms [see @scutari.denis2021] provide a means for searching through the (super-exponentially large) possible space of Bayesian networks to find the structure most likely to have generated some observed data. These approaches allow us to scale the Bayesian network modeling framework beyond the smaller models characteristic of early computational work on intuitive theories [e.g., @griffiths.tenenbaum2005].

@@

Across all our studies, our primary outcome of interest was participants’ intentions to vaccinate their children against diseases like measles, mumps, and rubella (which we refer to henceforth as *vaccination intentions*). Drawing on a variety of sources, including research on anti-vaccine skepticism [e.g.,  @dube.etal2015; @larson.etal2015; @salmon.etal2015; @williams2014], anti-vaccine websites, and our own qualitative work with vaccine skeptics (see Supplementary Materials), we generated a list of 13 additional beliefs that, though surely not exhaustive, might jointly offer a reasonable means of understanding and predicting vaccine intentions.

@@

Each scale measured agreement with 4-6 statements, including at least one reverse-coded item, and was highly reliable (observed reliability, Study 1: all Cronbach’s $\alpha \geq .73$). We developed 13 of these scales, and included one preexisting scale (the “holistic balance” subscale from [@mcfadden.etal2010]. See Supplementary Materials for all items included in all scales.

In Study 1, we administered these 14 belief scales to participants in an online survey with `r results$s1$demo$n_include` participants. We then used these intercorrelations to build a cognitive model of the relations among these beliefs via Bayesian network structure learning techniques.

@@

A total of `r results$s1$demo$n_total` people recruited from Amazon Mechanical Turk (mTurk) participated in this study conducted in August of 2017. Participants were paid \$1.60 for about 8 minutes of their time. `r results$s1$demo$n_exclude` participants (6%) failed at least one of two attention check questions and were excluded from further analysis, leaving a final sample of n = `r results$s1$demo$n_include`. See Supplementary Materials for demographic information about this sample.

@@

The goal of this study was to use these data to conduct a bottom-up search for a Bayesian network that would represent the intuitive theory surrounding vaccination decisions as a network of related beliefs. There were strong intercorrelations among these 14 beliefs confirming that these beliefs are, in fact, closely related to each other and to intentions to vaccinate children (see SM). Based on the observed associations among participants' beliefs about each of theses issues, we sought to infer how these states of affairs (e.g., that vaccines are dangerous, that diseases are severe) are connected in people's intuitive theories---how people understand these states of affairs to cause, imply, or otherwise influence one another.

@@

A Bayesian network is a directed acyclic graph that represents variables as nodes connected by directed edges. Each set of “child” and “parent” nodes in the larger network is a “family” (for some examples, see Figure 1b). The network’s directed edges encode conditional dependencies among variables, and many parameterizations for the individual families are possible. In our model, the probability of some “child” belief node $y$ can be expressed as a noisy-logical function of the influence of one or more “parent” belief nodes $x$:

\begin{equation}
P(y=1) = \Big(1-(1-w_0)\prod_i(1-w_i)^{x_i}\Big)\prod_j(1-w_j)^{x_j}
\end{equation}

@@

This parameterization is a substantial simplification compared with the full flexibility possible within a Bayesian network. A cognitive model composed of these types of relations is capable of capturing a causal system of non-interactive generative and preventive causes. But it bears noting that this simplification comes with limitations. For instance, returning to the example network in Figure 1c, this parameterization would not permit capturing the relationships between dryness, lightning, and fire. Dryness is an enabling condition that both increases the probability of fire but also increases the probability that lightning would cause a fire. So the influences of these variables are not independent and would need a more complex parameterization to be properly represented.

Finally, having transformed participants’ scores on each belief scale to fall between 0 and 1, we estimated the parameters of this model as a Beta regression using maximum likelihood methods (see Supplementary Materials for details).

@@

We aimed to represent not only the relationships among variables in the data but the *cognitive model* that gave rise to these observed relationships, i.e., the relationships among beliefs in the mind. We actually estimated two models sharing the same network structure as depicted in Figure 2: First, we estimated a data model assuming that participants possess a shared mental model from which they generate their responses to the belief scales. In this model, the values of $p(y=1)$ and $x$ in the equation above corresponded to participants’ continuous responses to the belief scales (e.g. a participant's score for the *vaccine danger* scale). Then, from the parameters estimated in the data model we created a cognitive model of participants’ reasoning, where the $x$ are binary, indicating whether that state of affairs obtains (e.g., whether “vaccines are dangerous”).

We used this cognitive model to simulate participants' degrees of beliefs given other beliefs, as well as how beliefs would change following exposure to evidence.

@@

To conduct an efficient search through the super-exponential space of graphs, we tested several learning algorithms implemented in the “bnlearn” R package [@scutari2010] and evaluated their performance through cross validation. To avoid overfitting, we used an 80% split of the data (*n* = 904) for training, and reserved the remaining 20% for evaluating model performance. We compared performance of different structure learning algorithms using k-fold cross validation within the training split. Graphs were scored by Bayesian Information Criteria, an asymptotic estimate of the likelihood of the data under a given graph structure. Based on these results, we chose to apply a score-based "hill-climbing"  algorithm using our generative model constraints (see Supplementary Materials for details). Applying these techniques resulted in an estimate of the network shown in Figure 1.

```{r fig1-maingraph, fig.align='center', out.width="90%", fig.cap="Bayesian network cognitive model resulting from application of hill-climbing algorithm with our generative-model constraints applied to data in Study 1 training split (n = 904). The network has 29 edges connecting these 14 variables. Edge polarity and weights are indicated by color, with red hues indicating negative or preventive influences, and blue hues indicating positive or generating influences. The direction of the edge was not determinable from the data and modeling constraints for two edges: $\\text{Naturalism} \\rightarrow \\text{Holistic Balance}$ and $\\text{Medical Skepticism} \\rightarrow \\text{Parental Expertise}$; see Supplementary Materials."}

@@

We evaluated the resulting cognitive model’s performance by testing its accuracy in predicting responses in the 20% testing split (*n* = 226). We generated predictions for held-out participants’ responses for each variable by conditioning the cognitive model network on the remaining 13 observed variables. Figure 3 compares the model’s predictions with participants’ actual responses. Collapsing across all variables, the average correlation between predicted and observed responses was *r* = `r round(avg_corr, 3)`, accounting for `r round(100*avg_corr^2, 0)`% of the variance in observed responses.

Our goal in developing this cognitive model was to create a psychologically plausible and intelligible model of the intuitive theory surrounding vaccination decisions. However, it is also reasonable to ask whether other less-constrained models might be more predictive. To address this, we compared the Bayesian network model’s predictions to a suite of 14 naive “maximal” regression models, each using values on all 13 other belief scales as predictors for each belief. The model presented in Fig. 1---which we take to be simpler, more psychologically plausible, and more intelligible---performed as strongly as these "maximal" regression models (see Supplementary Materials).

```{r fig2-validation, fig.align='center', out.width="100%", fig.cap="Observed responses and cognitive model predictions for each belief among participants in the Study 1 testing split. Correlations between observed and predicted values ranged from .380 to .898 across the different belief scales. In general, the model showed greater predictive accuracy for more central beliefs (e.g., \\emph{vaccine danger}) than for more peripheral beliefs (e.g., \\emph{parental protectiveness}; see Supplementary Materials for details)."}
knitr::include_graphics("../local/valid_plt.png")
```

#### What does this cognitive model reveal about the intuitive theory underlying vaccination decisions?

@@

It is worth pointing out that not all of the relationships represented in the model should be interpreted as _causal_. Due to Bayesian networks’ tight connections with structural causal networks, it can be tempting to give the edges this interpretation, but in some cases these relationships are better thought of as probabilistic dependencies or “influence” relations [@williamson2001]. For instance, the edge running from vaccine danger to toxic additives in vaccines clearly does not represent a causal influence: vaccines being dangerous does not cause them to contain toxins. Instead, a more plausible interpretation of this edge is that it reflects something like a set membership relation: that toxins are one type of danger that might be present in vaccines.

In other cases, our data and explicit constraints do not perfectly identify the directions of the edges. This is the case for two edges (Naturalism $\rightarrow$  Holistic Balance and Medical Skepticism $\rightarrow$ Parental Expertise; see Supplementary Materials for details). In both cases, we do not feel it is intuitive which direction the edge should flow in, nor even that it should be directed at all. Such a counterintuitive result might be taken as indicating a limitation of Bayesian network models, which are composed solely of directed, acyclic relationships. We note that prior work has successfully modeled coherent belief revision by leveraging models with undirected or bidirected relations [e.g. @holyoak.simon1999; @simon.etal2004; @thagard1989; @spellman.etal1993]. Alternatively, we might instead hypothesize there is an additional unmeasured variable that connects these beliefs that could more intelligibly account for their association within the context of a directed network. For instance, a more abstract epistemic belief might influence both medical skepticism and parental expertise beliefs. To speculate, such an abstract belief might be in one's own expertise or in how we come to have knowledge (e.g. through direct experience versus science).

More generally, as they operate solely on correlations among variables, structure learning algorithms cannot be guaranteed to correctly identify causal directionality. Looking forward to future applications of these methods, researchers might seek other ways to resolve ambiguities left by structure learning algorithms. The simplest approach would be to set edges manually based on top-down theorizing, or even to set stronger constraints in advance of fitting models. Another potential future direction would be to probe people's mental causal models more directly, either by asking them directly what they think about the causal relationships among variables or by asking them to reason through hypothetical interventions on variables. These more direct responses might then be used to constrain or augment the structure learning algorithm's outputs.

@@

Very few educational interventions have been shown to improve attitudes toward childhood vaccination. One of the few successful educational interventions in this domain was tested by Horne and colleagues [-@horne.etal2015], who found that emphasizing the dangers of preventable diseases like mumps, measles, and rubella improved participants’ attitudes toward vaccines. Their intervention included descriptions of the symptoms of mumps, measles, and rubella; vivid photos of children suffering from these diseases; and a brief story from the perspective of a parent whose child was hospitalized with measles (see Supplementary Materials). This “disease risk” intervention improved attitudes toward vaccines compared to an irrelevant control essay and compared to an intervention aimed at debunking concerns about vaccines causing autism. In Study 2, we aimed to replicate this success and leverage the cognitive model developed in Study 1 to gain insight into how and why such an intervention successfully induces  belief revision in this domain.

In Study 2, we used the cognitive model established in Study 1 to predict how Horne and colleagues’ [-@horne.etal2015] intervention would affect the beliefs in this network, including people’s intentions to vaccinate their children. Participants completed the 14 belief scales from Study 1, then returned two days later to complete either the disease risk intervention or a no-intervention control, followed by the same 14 belief scales.

@@

Participants (*n* = `r results$s2$demo$n_complete` passed attention checks in total) in the intervention conditions were presented with a version of Horne and colleagues'  “disease risk” intervention, modified to present subparts in a fixed (rather than randomized) order, with slight changes to wording to improve clarity, and with a final statement summarizing the take-home message of the intervention; see Supplementary Materials. They then completed the same 14 belief scales and demographic questions that they had completed in Phase 1.

@@

Figure 4 shows the change in participants’ attitudes across all 14 belief scales, from pretest to posttest. As expected, the largest change was observed for *disease severity* beliefs---the belief directly targeted by the intervention. Most importantly, the disease risk intervention significantly increased participant’s *vaccination intentions*, conceptually replicating earlier findings regarding this intervention [@horne.etal2015].

@@

To model the effects of the intervention presented in Study 2 on the intuitive theory surfaced in Study 1, we augmented the cognitive model's network representing the intuitive theory with an additional node representing the information about measles, mumps, and rubella presented in the intervention. This evidence node was added as a child of the *disease severity* node, allowing the intervention to act as virtual evidence for *disease severity* [following @pearl1988].

@@

\begin{equation}
logit(posttest) = logit(pretest) + \ln(ER_0) + x\ln(ER_{1})
\end{equation}

@@

\begin{equation}
ER_i = \ln\frac{(P(d_i|h=1))}{(P(d_i|h=0))}
\end{equation}
